### 2012-09-20 doc/prelim/document/chapters/linear_problem/linear_problem.tex
adding example from gentile paper

\subsubsection{Domain Decomposition for Monte Carlo}
\label{subsubsec:msod}
As observed in the discussion on parallel Krylov methods, large-scale
problems will surely have their data partitioned such that each
parllel process owns a subset of the equations in the linear
system. Given this convention, the adjoint Monte Carlo algorithm must
perform random walks over a domain that is decomposed and must remain
decomposed due to memory limitations. This naturally leads us to seek
parallel algorithms that handle domain decomposition. In the context
of radiation transport, Brunner and colleagues provided one of the
first surveys of algorithms for achieving this
\citep{brunner_comparison_2006}. In their work they identity two data
sets that are required to be communicated: the sharing of particles
that are transported from one domain to another and therefore from one
processor to another and a global communication that signals if
particle transport has been completed on all processors. The
algorithms presented include a fully-locking synchronous scheme, a
fully-asynchronous pattern, a traditional master/slave scheme, and a
modified master/slave scheme the implements a binary tree pattern for
the global reduction type operations needed to communicate between the
master and slave processes.

In 2010, Wagner and colleagues developed the \textit{multiple-set
  overlapping-domain} (MSOD) decomposition for parallel Monte Carlo
applications for full-core reactor analysis
\citep{wagner_hybrid_2010}.

>>>>

\subsubsection{Domain Decomposition for Monte Carlo}
\label{subsubsec:msod}
As observed in the discussion on parallel Krylov methods, large-scale
problems will surely have their data partitioned such that each
parllel process owns a subset of the equations in the linear
system. Given this convention, the adjoint Monte Carlo algorithm must
perform random walks over a domain that is decomposed and must remain
decomposed due to memory limitations. This naturally leads us to seek
parallel algorithms that handle domain decomposition. In the context
of radiation transport, Brunner and colleagues provided a survey of
algorithms for achieving this as implemented in production implicit
Monte Carlo codes \citep{brunner_comparison_2006}. In their work they
identity two data sets that are required to be communicated: the
sharing of particles that are transported from one domain to another
and therefore from one processor to another and a global communication
that signals if particle transport has been completed on all
processors. The algorithms presented are a fully-locking synchronous
scheme, a asynchronous-send/synchronous-receive pattern, a traditional
master/slave scheme, and a modified master/slave scheme the implements
a binary tree pattern for the global reduction type operations needed
to communicate between the master and slave processes. They observed
that the modified master/slave scheme performed best in that global
communications were implemented more efficiently than those required
by the asynchronous scheme. Furthermore, none of these schemes handled
load-imbalanced cases efficiently. Such cases will be common if the
source sampled in the Monte Carlo random walk is not isotropic and
evenly distributed throughout the global domain. It was noted that
efficiencies were improved by increasing the frequency by which
particle data was communicated between domain-adjacent processors.
However, this ultimately increases communication costs.

The 2006 work of Brunner is notable in that the Monte Carlo codes used
to implement and test the algorithms adhered to a strict policy of
generated identical results, indepdent of domain decomposition or
domain replication as derived from the work of Gentile and colleagues
\citep{gentile_obtaining_2005}. In their work, a procedure is given for
obtaining results reproducable to machine precision for an arbitrary
number of processors and domains. Differences can arise from using a 
different random number sequence in each domain and performing
a sequence of floating point operations on identical data in a
different, leading to variations in round-off error and ultimately a
non-identical answer. They use a simple example, recreated below in
Figure~\ref{fig:gentile_example}, that illustrates these issues.
\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/linear_problem/gentile_example.pdftex_t} }
  \end{center}
  \caption{\textbf{Gentile's example illustrating how domain
      decomposition can create reproducability issues in Monte Carlo.}
    \textit{Both particles A and B start in zone 1 on processor
      1. Particle A moves to zone 2 on processor 2 and scatters back
      to zone 1 while B scatters in zone 1 and remains there.}}
  \label{fig:gentile_example}
\end{figure}

In 2010, Wagner and colleagues developed the \textit{multiple-set
  overlapping-domain} (MSOD) decomposition for parallel Monte Carlo
applications for full-core reactor analysis
\citep{wagner_hybrid_2010}. In their work, a certain amount of overlap
between domains was used to decrease the number of particles leaving
the domain.
